Wireless service accounting based on Ethereum state channels - Master Thesis

### 2019-10-25 Paper.tex
Add risk analysis, correct some things

\begin{abstract}
  \begin{center}
    This document contains authors' ``\textit{Blockchain
      technologies master}''\footnote{UPC School
      of Professional \& Executive Development}
    final thesis. The project described in the document
    consists on a technological proof
    of concept of an internet access service consumption
    using a state channel smart contract as
    mean of payment. After a brief
    introduction, there's a first motivation of
    the use case included;
    in the following sections
    there's a detailed explanation of
    the architecture, implementation and
    security considerations;
    after that it is described a more
    general approach about using blockchain technologies
    and state channels in particular to provide
    mechanisms to reduce advanced wireless
    (i. e. 5G) services market friction between neutral host
    providers, virtual operators and operators
    purchasing licenses of spectrum bands to nations' governments;
    guarantee transparency to consumers,
    and provide a fully GDPR compliant
    schema relying on a self-sovereign identity system.
  \end{center}
\end{abstract}

@@

\begin{center}
\Huge{\textbf{DISCLAMER}}
\vspace{1cm}

\Large
All software described and linked in this document is open source
under the MIT license; so anyone can freely use it and make
modifications to the source code to fulfill their use
cases. But if you do so, please take into account that the software
at the time of this writing is merely
a proof of concept and is not properly audited; although
authors tried to follow best practices regarding smart contract,
front-end and back-end design and implementation there are still
some tradeoffs and recognizable security issues
described in this document's
security considerations section. Please take this into account
when using and/or modifying this software and be conscious
that authors do not provide any guarantee in case of
security issues, economical loses or regulation infringement.
\end{center}

@@

Taking this into account, lead the authors' to develop
a first bottom-up proof of concept of a simple system
demonstrating how can a wireless WiFi hotspot
deployed on a commodity hardware asset\cite{RaspberryPi19}
be managed to allow users to access the internet
paying in exchange for Ethereum on a public network.

@@

One of the first intiatives to create
a bandwidth reselling market was La Fonera\cite{fon}
back in 2006. Main goal of the initiative was
to allow ADSL users to share unused bandwidth using
a specific WiFi hotspot deployed on an ad-hoc
built router.

@@

A more recent project like Hotspot Me\cite{hotspotme}
and also very aligned with the technological demonstration
described in this document consist on using mobile
APIs to allow users to share WiFi hotspot tethering
to users nearby in exchange for ETH supported by a
micro-payment channel contract.

@@

The consensus mechanism ensures a common order
of unambiguous transactions and
blocks, and guarantees integrity and
blockchain consistency a
through geographically distributed nodes.
By its design, the blockchain has characteristics
such as: decentralization, integrity and auditability.
The blockchain can serve as a new
type of software connector, which should be considered
as a possible decentralized alternative to storage
of existing centralized shared data.

@@

\begin{itemize}
\item Is the most used and mature
turing-complete-programmable blockchain technology.
\item Support for high quality and audited
libraries in Solidity (most used smart contract language in Ethereum).
\item Developments on Ethereum can be deployed
over public (main or testnets) or private/permissioned networks.
\item Unlike Bitcoin, Ethereum supports miner fees using gas
instead of directly ETH; this can mitigate the effects of
volatile crypto-currency price by adjusting gas price depending
on ETH value.
\end{itemize}

@@

In order to support the PoC a
Raspberry Pi 3 Model B (see figure 2) has been used
as a single board computer to host all the required software.
Tested operative system was OSMC\cite{osmc} (a Debian based
GNU-Linux flavor) mainly used as media-center.
A WiFi access point using
hostapd\cite{hostapd} daemon over Raspberry's WiFi chipset.
Raspberry Pi was also connected using ethernet cable
to an optical fiber commercial router (\textit{see figure 2}).

@@

% \begin{tabular} to 0.5\textwidth { | X[l] | X[r] | }
\begin{tabular}{ | m{0.125\textwidth} | m{0.3125\textwidth} | }
 \hline
 Specs & Raspberry Pi 3 Model B \\
 \hline
 CPU & Broadcom BCM2837, Cortex-A53 (ARMv8) 64-bit SoC @ 1,2 GHz \\
\hline
 RAM & 1 GB \\
\hline
 Connectivity & Wi-Fi 802.11 b/g/n (2,4 GHz) Bluetooth 4.1 Ethernet card up to 100 Mbps \\
\hline
 Ports & Full HDMI, 4 USB 2.0, MicroSD, CSI camera, DSI display \\
\hline
 Memory & MicroSD \\
\hline
\end{tabular}
\end{center}
\vspace{0.15cm}
Table 1. Raspberry Pi 3 Model B specs
Although it is not expressly indicated if it is free hardware
(\textit{open hardware}) or with trademark rights, on the
official website there's an explanation regarding
that Raspberry have distribution and sales contracts with
two companies and guarantee that anyone can become a reseller or
redistributor of Raspberry Pi cards [RaspberryPiBuy 19].

@@

Main idea is that the users can join freely
to the wireless hotspot and access the web captive portal
(\textit{hosted on the same Raspberry}) application that
communicates with a back-end that controls iptables
rules and also an Infura light proxy able to relay RPC
calls arriving on 8545 port of the Raspberry.

@@

\begin{itemize}
\item[] 1) The user associates the laptop with the WiFi hotspot.
\item[] 2) Once connected is redirected to the captive portal front-end.
\item[] 3) Front provisions MetaMask as Web3 provider.
\item[] 4) Front-end uses MetaMask to allow the user
  to send value to the contract
\item[] 5-6) Signed transaction is relayed through a light Infura proxy listening
  on port 8545.
\item[] 7) Once transaction is confirmed, the back-end iptables controller
  start forwarding user's traffic.
\end{itemize}

\subsubsection{Channel transactions}
Before sending the initial value transaction (\textit{proportional to
  the initially expected duration of the service})
user's front-end generates an ephemeral elliptic curve
secp256k1 keypair and sends the public key to the contract to open the state channel.
Corresponding ephemeral private key is the one that's going to be used
to periodically (each 10 s approximately)
produce increasing value signatures proportional to the
service consumption ellapsed time. This signatures are being send to the
backend instead of the blockchain; backend verifies and validates the
signatures and stores the last one (\textit{the most valuable}).

\subsubsection{Channel closing}
At the end, when the connection time is exhausted or the backend
stops receiving channel transactions for approximately 60 seconds,
the backend sends the last seen transaction to the smart contract
using the method \texttt{closeChannel}; this method, when called
with a valid channel signature as parameter, unlocks user's original
funds and transfer the signed amount to the contract owner
(\textit{provider of the service}) and the corresponding return
to the user.

@@

This section includes a detailed rationale software description of all
the components and subcomponents.

@@

Main data stucture of the contract described below:

@@

\paragraph{\texttt{ChannelData}} Includes payer ethereum address,
its state channel ephemeral address, deposit amount, opening time
and a boolean representing the open/close status of the channel.

@@

\paragraph{\texttt{channelCount}} uint256 variable that increments
in a unit every time a user opens a channel. This operation
does not rely on SafeMath since it's virtually impossible to
overflow a uint256 storage record.

@@

\paragraph{\texttt{openChannel}} receives as parameters, the ephemeral address
of the user and also the purchased service value.

@@

\paragraph{\texttt{closeChannel}} accepts the amount of ETH to refund
to sender, the \texttt{channelId} and the \texttt{signature} of the
payer; checks the signature and after updating the state, transfers
the amount to the owner of the contract and the return to the channel
opener.

@@

\paragraph{\texttt{claimTimeout}} method that allows the payer to
close the channel in case contract owner does not close it after
some \texttt{expirationTime}.

@@

Consists on a NodeJS backend that fulfils three main system functions:

@@

On execution uses the module \texttt{lib/iptables.js}
to setup and initialize system's iptables redirecting
all 80 port traffic to the ip address to Raspberry's
ip address; to do so, relyes on the node module \texttt{child\_process}
to spawn system commands. Also redirects all traffic trying
to reach 8545 port to itself the same way in order to
redirect MetaMask's RCP calls to the lightweight infuraProxy\cite{infuraProxy}
developed also for this project and contained in etherwall's root directory.

@@

After initializing, the backend using the web framework\cite{express}
etherwall exposes the following endpoints:

@@

\paragraph{\texttt{POST /mac}} Endpoint to request a new
connection; mainly
accepts the time left to use the service as a parameter
and internally gets the MAC address of user's device
making an arp table query based on client's ip address.
After that invokes \texttt{lib/iptables.js} \texttt{grantAccess}
function in order to insert a rule to enable packet forwarding
for this particular MAC address on top of iptables.
Also uses module \texttt{lib/connections.js} to keep persistence
of the connection status. Once connected after posting
\texttt{timeLeft} to this endpoint, in module
\texttt{lib/connections.js} there's a \texttt{setInterval}
triggering getting connection status and time since last
micro-payment each 5 seconds. In case \texttt{timeSinceLastPayment}
exceeds 60 seconds or \texttt{timeLeft} goes negative,
disconnects the client deleting the forwarding rule from
iptables.

@@

\paragraph{\texttt{POST /payment}} Endpoint that allows the front-end
to send a payload containing a payload passing the corresponding
contract data:

@@

In order to verify the correctness of the payload, uses Web3
module to compute the \texttt{soliditySha3} and compares it
to the signature's \texttt{messageHash}; after that recovers
the public address from the signature and compares it to
the associated smart contract mapping structure based on its
\texttt{channelId}; if everything matches, it additionally checks
that the unlocked amount is approximately what it corresponds
taking into account connection ellapsed time
(\textit{using a kind threshold of 60 seconds marginal cost});
in case everything is correct, stores the payment together
with a timestamp as the last one corresponding
to the connection using again \texttt{lib/connections.js} module.
This is done to allow bootstrapped \texttt{setInterval}
on the previously described \texttt{POST mac/} to process
the timeout business logic.

\subsubsection{etherdoor} A very simple front-end based on React
using material styling; App's main \texttt{componentDidMount}
initializes the state including Web3 object and smart contract
wrapper, gets current connection status from backend's
\texttt{GET mac} endpoint and creates the micro-payment
state channel ephemeral account; at the end triggers a
\texttt{setInterval} loop that when connection activates,
will be responsible of sending proper signed micro-payments.

\begin{center}
  \includegraphics[keepaspectratio, width=0.4\textwidth]{images/portal-y.eps}
  \\
  Figure . \textit{etherdoor front-end landing page}
\end{center}

Usage depends on some button clicking, triggering
\texttt{someMinutesPass} function invokation that launches MetaMask
as Web3 provider making possible to the user to open a channel
on the smart contract building a transaction and relaying it
through \texttt{infuraProxy}; it also invokes \texttt{startChannel}
that will update React's App state and will cause
\texttt{signMicroPayment} function to periodically on a 10 seconds
interval to sign and post to the backend the increasing amount
micro-payments.

@@

To validate the PoC, a battery of tests has
been defined. For example, given
that smart contract normally handle money,
it is essential to ensure that
its number of failures and vulnerabilities is low [Hegedus 18].

@@

During development and in order to ease
the automatic deployment of the contract
over a Truffle's suite Ganache-CLI\cite{ganache}
instance there are two util scripts on the
repository\cite{state-channel-contract-a}
\texttt{utils/compile.js} and \texttt{utils/deploy.js}
that take care of merging all the Solidity source
code required and deploying it to the Ganache instance.

@@

Non-trivial tests covering main workflows include:
\begin{itemize}
\item Contract requires a minimum value amount to open a channel.
\item A user can open a channel sending value over a minimum amount sending
  also an Ethereum ephemeral public key
\item Prevention of closing channel operation using a different private key
  in signature.
\item Ability of the contract owner to close the channel using a properly
  generated user signature.
\item If channel is not closed by the owner, the user, cannot close it
  before a certain amount of time.
\item In case reaching expiration time without the owner properly closing the channel,
  the user can close the channel receiving in return unexpended amount.
\end{itemize}

@@

Both components were heavily manually tested, during development
in order to assure the correct workflows were correctly supported.

@@

To help developers and make technology more mature,
we need analysis tools [ConsenSys 19].

@@

For this project, we used Mythril from ConsenSys. Mythril is a
security analysis tool for EVM (Ethereum Virtual Machine) bytecode.
It detects security vulnerabilities in smart contracts built for Ethereum.

@@

We analysed the Solidity-based smart contract called
StateChannel.sol (of this project).

@@

https://github.com/ethereum-internet-access/mythril
Firstly, we analysed the local source code:

@@

Secondly, we analysed on-chain smart contract.
Because the smart contract was deployed on Ropsten testnet we executed:

@@

We simply install Docker and run the container over Raspbian/ARM.
It would deploy the done project. The script to execute is as follows:

@@

\begin{itemize}
\item Add firewall rules to prevent IP traffic between
  different end users.
\item Obviously securize the connectivity to the Raspberry
  using a wire connection and filtering any traffic on 22 (ssh)
  port coming from the wireless interface.
\item For this PoC, using only the captive portal IP address; TLS
  is not supported; accordingly the user has no guarantee of
  authenticity of the WiFi service provider.
\end{itemize}

@@

All smart contract-blockchain transactions are being processed
using Infura as an Ethereum proxy; only security issue regarding
private key management arises when considering the ephemeral
private keypair generation on the front-end.

@@

@@

Any organisation that has customers on the premises,
from coffee shops to libraries, might want to offer free (or even
paid-for) Wi-Fi as an attraction. Collecting and storing personal
information brings responsibilities under European Union’s General
Data Protection Directive (GDPR). The GDPR\cite{gdpr16}, which took
effect on 25 May 2018, applies to any organisation that handles EU
residents’ personal data. From 25 May on, however, organisations
failing to comply with the regulation face fines of up to 4 \% of
their global turnover. There are Certified EU General Data Protection
Regulation Foundation (GDPR) Training Courses.

@@

We take another look at the major obligations imposed by this
regulation\cite{tata17}:

\vspace{0.35cm}
1. ART 5/1B and 1C - DATA MINIMISATION:
Prohibited is the collection of more data than that required for the
adequate purposes for which it is being processed. For instance, if
you offer restaurant guests free WiFi access it is not necessary to
ask for their address data in this connection. However, if you are
interested in finding out where your guests are from - which can be a
completely legitimate interest for a restaurant owner to have - you
are indeed entitled to pose the question, but your guests must divulge
the information consciously and voluntarily and perhaps you should
only ask them for their postcode.

@@

\vspace{0.35cm}
5. ART 5/1F and 28 - INTEGRITY AND CONFIDENTIALITY: You must ensure
appropriate security of data and also ensure that it cannot be
tampered with. But how do you do that? What if hackers launch a
malicious attack? Don’t worry; even government websites are not immune
to this sort of thing and it would be hardly fair if the GDPR were to
impose more stringent requirements on you. However, you are required
to put appropriate technical or organisational measures in place that
provide adequate protection of your customers’ data. The GDPR does not
spell out in so many words what these required ‘measures’
are. However, you can be certain that you will be at least be expected
to have a valid SSL certificate for your data-collecting website and
to conclude instruction and confidentiality agreements with your
personnel who come into contact with the corresponding data.

@@

In addition, the Proposal for a new ePrivacy
Regulation was tabled in January 2017, when GDPR was not yet
applicable and the European Electronic Communications Code\cite{eecc18}
was still being negotiated. Now, GDPR has been implemented,
changing many of the approaches and behaviours that the ePrivacy
Regulation aims to tackle. Additionally, the new EECC already ensures
an extended definition of what an e-communications service is, thereby
achieving one of the purported goals of the ePrivacy Regulation, which
is to broaden the scope of application of confidentiality obligations
to a larger number of Internet-based players.

\vspace{0.35cm} Because of these issues, good advice for anyone
planning to set up a hotspot would be to collect as little personal
data from users as possible, and to think of any additional collection
as a potential liability.

@@

In the case of 5G technologies for practical reasons
and the high density of antennas required,
different operators shall be able
to dynamically share the same resources (i.e. neutral host
infrastructure). At the same time in order to manage
service operation efficiently through its whole
life-cycle there's the need to automate as much as
possible all the operations; this requirements
perfectly fit smart contract capabilities, in order
to model and operate over a defined set of rules
supporting payments, escrows, rewards or penalties
depending on different conditions.

\end{multicols}
\begin{center}
%\begin{figure}[h]
  %  \centering
  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/slices-y.eps}
%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
%\end{figure}
\\
Figure 5. Slicing model.
\\
\end{center}
\begin{multicols}{2}

As an example, in figure 5 it's shown how three virtual operators over
different infrastructures shared or not in a slice-model of a neutral host
infrastructure; a slice can be described as a dynamic virtualization
allowing multitenancy of some physical resources like access points,
physical switches and datacenters including mobile edge computing
appliances; this is the case of currently ongoing projects like
5GCity\cite{5gcity} where a distributed cloud
and radio platform for municipalities and infrastructure owners acting
as 5G neutral hosts is being developed and tested. Platforms
targeting 5G business models such as 5GCity can benefit from adding
this smart contract business logic in order to bill end-users and automate
all the accounting layer.

@@

Although this is a promising approach; there's an issue with it regarding
regulation, in particular in the European Union, there's the requirement
that a telecommunication service provider must know their customers; this
is required for example to allow judge requirements of users identity
in case there's some evidence of illegal activities. As a single payment
system, state channel smart contracts are pseudoanymous hence do not
support the \textit{Know Your Customer - KyC} required by regulation.

@@

To take into account the problem illustrated in previous paragraph
we can take as a basis a model of self-sovereign identity\cite{allen}
and in particular the one recently anounced by the authorities
of Catalonian government\cite{identicat} that will provide credentials
to citizens using DLT technologies as a Public Key Infrastructure; this
model could allow in some near future to give the citizen the option
of registering identity in such a way that the credential can be linked
to some administration database that could be queried only by a judge
requirement and guarantee the users identities are recognized

@@

\end{thebibliography}
\clearpage

>>>>

\begin{abstract}
  \begin{center}
    This document contains authors' ``\textit{Blockchain
      technologies master}''\footnote{UPC School
      of Professional \& Executive Development}
    final thesis. The project described in the document
    consists on a technological proof
    of concept of an internet access service consumption
    using a state channel smart contract as
    mean of payment. After a brief
    introduction, there's a first motivation of
    the use case included;
    in the following sections
    there's a detailed explanation of
    the architecture, implementation and
    security considerations;
    after that it is described a more
    general approach about using blockchain technologies
    and state channels to provide
    mechanisms to reduce advanced wireless
    (i. e. 5G) services market friction between neutral host
    providers, virtual operators and operators
    purchasing licenses of spectrum bands to nations' governments;
    ideally achieving an unprecedented transparency guarantee to consumers and
    a fully GDPR compliant
    schema relying on a self-sovereign identity system.
  \end{center}
\end{abstract}

@@

\begin{center}
\Huge{\textbf{DISCLAIMER}}
\vspace{1cm}

\Large
All software described and linked in this document is open source
under the MIT license; so anyone can freely use it and make
modifications to the source code to fulfill their use
cases. But if you do so, please take into account that the software
at the time of this writing is merely
a proof of concept and is not properly audited; although
authors tried to follow best practices regarding smart contract,
front-end and back-end design and implementation there are still
some trade-offs and recognizable security issues
described in this document's
security considerations section. Please take this into account
when using and/or modifying this software and be conscious
that authors do not provide any guarantee in case of
security issues, economical loses or regulation infringement.
\end{center}

@@

Taking this into account, lead the authors' to develop
a first bottom-up proof of concept of a simple system
demonstrating how can a wireless WiFi hot-spot
deployed on a commodity hardware asset\cite{RaspberryPi19}
be managed to allow users to access the internet
paying in exchange for Ethereum on a public network.

@@

One of the first initiatives to create
a bandwidth reselling market was La Fonera\cite{fon}
back in 2006. Main goal of the initiative was
to allow ADSL users to share unused bandwidth using
a specific WiFi hot-spot deployed on an ad hoc
built router.

@@

A more recent project like Hot-spot Me\cite{hotspotme}
and also very aligned with the technological demonstration
described in this document consist on using mobile
APIs to allow users to share WiFi hot-spot tethering
to users nearby in exchange for ETH supported by a
micro-payment channel contract.

@@

The consensus mechanism ensures a common order
of unambiguous transactions and
blocks, and guarantees integrity and
blockchain consistency a
through geographically distributed nodes.
By its design, the blockchain has characteristics
such as: decentralization, integrity and audibility.
The blockchain can serve as a new
type of software connector, which should be considered
as a possible decentralized alternative to storage
of existing centralized shared data.

@@

\begin{itemize}
\item Is the most used and mature
Turing-complete-programmable blockchain technology.
\item Support for high quality and audited
libraries in Solidity (most used smart contract language in Ethereum).
\item Developments on Ethereum can be deployed
over public (main or test-nets) or private/permissioned networks.
\item Unlike Bitcoin, Ethereum supports miner fees using gas
instead of directly ETH; this can mitigate the effects of
volatile crypto-currency price by adjusting gas price depending
on ETH value.
\end{itemize}

@@

In order to support the PoC a
Raspberry Pi 3 Model B (see figure 2) has been used
as a single board computer to host all the required software.
Tested operative system was OSMC\cite{osmc} (a Debian based
GNU-Linux flavor) mainly used as media-center.
A WiFi access point using
hostapd\cite{hostapd} daemon over Raspberry's WiFi chip-set.
Raspberry Pi was also connected using ethernet cable
to an optical fiber commercial router (\textit{see figure 2}).

@@

% \begin{tabular} to 0.5\textwidth { | X[l] | X[r] | }
\begin{tabular}{ | m{0.125\textwidth} | m{0.3125\textwidth} | }
 \hline
 Specs & Raspberry Pi 3 Model B \\
 \hline
 CPU & Broadcom BCM2837, Cortex-A53 (ARMv8) 64-bit SoC @ 1,2 GHz \\
\hline
 RAM & 1 GB \\
\hline
 Connectivity & Wi-Fi 802.11 b/g/n (2,4 GHz) Bluetooth 4.1 Ethernet card up to 100 Mbps \\
\hline
 Ports & Full HDMI, 4 USB 2.0, MicroSD, CSI camera, DSI display \\
\hline
 Memory & MicroSD \\
\hline
\end{tabular}
\end{center}
\vspace{0.15cm}
Table 1. Raspberry Pi 3 Model B specs
Although it is not expressly indicated if it is free hardware
(\textit{open hardware}) or with trademark rights, on the
official website there's an explanation regarding
that Raspberry have distribution and sales contracts with
two companies and guarantee that anyone can become a re-seller or
re-distributor of Raspberry Pi cards [RaspberryPiBuy 19].

@@

Main idea is that the users can join freely
to the wireless hot-spot and access the web captive portal
(\textit{hosted on the same Raspberry}) application that
communicates with a back-end that controls iptables
rules and also an Infura light proxy able to relay RPC
calls arriving on 8545 port of the Raspberry.

@@

\begin{itemize}
\item[] 1) The user associates the laptop with the WiFi hot-spot.
\item[] 2) Once connected is redirected to the captive portal front-end.
\item[] 3) Front provisions MetaMask as Web3 provider.
\item[] 4) Front-end uses MetaMask to allow the user
  to send value to the contract
\item[] 5-6) Signed transaction is relayed through a light Infura proxy listening
  on port 8545.
\item[] 7) Once transaction is confirmed, the back-end iptables controller
  start forwarding user's traffic.
\end{itemize}

\subsubsection{Channel transactions}
Before sending the initial value transaction (\textit{proportional to
  the initially expected duration of the service})
user's front-end generates an ephemeral elliptic curve
secp256k1 keypair and sends the public key to the contract to open the state channel.
Corresponding ephemeral private key is the one that's going to be used
to periodically (each 10 s approximately)
produce increasing value signatures proportional to the
service consumption elapsed time. This signatures are being send to the
back-end instead of the blockchain; back-end verifies and validates the
signatures and stores the last one (\textit{the most valuable}).

\subsubsection{Channel closing}
At the end, when the connection time is exhausted or the back-end
stops receiving channel transactions for approximately 60 seconds,
the back-end sends the last seen transaction to the smart contract
using the method \texttt{closeChannel}; this method, when called
with a valid channel signature as parameter, unlocks user's original
funds and transfer the signed amount to the contract owner
(\textit{provider of the service}) and the corresponding return
to the user.

@@

This section includes a detailed rationale software description of all
the components and sub-components.

@@

Main data structure of the contract described below:

@@

\paragraph{\texttt{ChannelData}} Includes payer Ethereum address,
its state channel ephemeral address, deposit amount, opening time
and a boolean representing the open/close status of the channel.

@@

\paragraph{\texttt{channelCount}} uint256 variable that increments
in a unit every time a user opens a channel. This operation
does not rely on SafeMath since it's virtually impossible to
overflow an incremental
uint256 storage record according to transaction fees or block time delays.

@@

\paragraph{\texttt{openChannel}} Receives as parameters, the ephemeral address
of the user and also the purchased service value.

@@

\paragraph{\texttt{closeChannel}} Accepts the amount of ETH to refund
to sender, the \texttt{channelId} and the \texttt{signature} of the
payer; checks the signature and after updating the state, transfers
the amount to the owner of the contract and the return to the channel
opener.

@@

\paragraph{\texttt{claimTimeout}} Method that allows the payer to
close the channel in case contract owner does not close it after
some \texttt{expirationTime}.

@@

Consists on a NodeJS back-end that fulfills three main system functions:

@@

On execution uses the module \texttt{lib/iptables.js}
to setup and initialize system's iptables redirecting
all 80 port traffic to the ip address of Raspberry's
ip address; to do so, relies on the node module \texttt{child\_process}
to spawn system commands. Also redirects all traffic trying
to reach 8545 port to itself the same way in order to
redirect MetaMask's RCP calls to the lightweight infuraProxy\cite{infuraProxy}
developed also for this project and contained in etherwall's root directory.

@@

After initializing, the back-end using the web framework\cite{express}
etherwall exposes the following endpoints:

@@

\paragraph{\texttt{POST /mac}} Endpoint to request a new
connection; mainly
accepts the time left to use the service as a parameter
and internally gets the MAC address of user's device
making an arp table query based on client's ip address.
After that invokes \texttt{lib/iptables.js} \texttt{grantAccess}
function in order to insert a rule to enable packet forwarding
for this particular MAC address on top of iptables.
Also uses module \texttt{lib/connections.js} to keep persistence
of the connection status. Once connected after posting
\texttt{timeLeft} to this endpoint, in module
\texttt{lib/connections.js} there's a \texttt{setInterval}
triggering getting connection status and time since last
micro-payment on a 5 seconds frequency. In case \texttt{timeSinceLastPayment}
exceeds 60 seconds or \texttt{timeLeft} goes negative,
disconnects the client deleting the forwarding rule from
iptables.

@@

\paragraph{\texttt{POST /payment}} Endpoint that allows the front-end
to send a payload containing a JSON body passing the corresponding
contract data:

@@

In order to verify the correctness of the payload, uses Web3
module to compute the \texttt{soliditySha3} and compares it
to the signature's \texttt{messageHash}; after that recovers
the public address from the signature and compares it to
the associated smart contract mapping structure based on its
\texttt{channelId}; if everything matches, it additionally checks
that the unlocked amount is approximately what it corresponds
taking into account connection elapsed time
(\textit{using a kind threshold of 60 seconds marginal cost});
in case everything is correct, stores the payment together
with a time-stamp as the last one corresponding
to the connection using again \texttt{lib/connections.js} module.
This is done to allow bootstrapped \texttt{setInterval}
on the previously described \texttt{POST mac/} to process
the timeout business logic.

\subsubsection{etherdoor} A very simple front-end based on React
using material styling; App's main \texttt{componentDidMount}
initializes the state including Web3 object and smart contract
wrapper, gets current connection status from back-end's
\texttt{GET mac} endpoint and creates the micro-payment
state channel ephemeral account; at the end triggers a
\texttt{setInterval} loop that when connection activates,
will be responsible of sending proper signed micro-payments.

\begin{center}
  \includegraphics[keepaspectratio, width=0.4\textwidth]{images/portal-y.eps}
  \\
  Figure 5. \textit{Etherdoor front-end landing page}
\end{center}

Usage depends on some button clicking, triggering
\texttt{someMinutesPass} function invocation that launches MetaMask
as Web3 provider making possible to the user to open a channel
on the smart contract building a transaction and relaying it
through \texttt{infuraProxy}; it also invokes \texttt{startChannel}
that will update React's App state and will cause
\texttt{signMicroPayment} function to periodically on a 10 seconds
interval to sign and post to the back-end the increasing amount
micro-payments.

@@

To validate the PoC, a battery of tests has
been defined. For example, given
that smart contract normally handle money,
it is essential to ensure that
its number of failures and vulnerabilities is low\cite{Hegedus18}.

@@

During development and in order to ease
the automatic deployment of the contract
over a Truffle's suite Ganache-CLI\cite{ganache}
instance there are two until scripts on the
repository\cite{state-channel-contract-a}
\texttt{utils/compile.js} and \texttt{utils/deploy.js}
that take care of merging all the Solidity source
code required and deploying it to the Ganache instance.

@@

Non-trivial tests covering main work-flows include:
\begin{itemize}
\item Contract requires a minimum value amount to open a channel.
\item A user can open a channel sending value over a minimum amount providing
  in addition an Ethereum ephemeral public key
\item Prevention of channel closing operation using a different private key
  in signature.
\item Ability of the contract owner to close the channel using a properly
  generated user signature.
\item If channel is not closed by the owner, the user, cannot close it
  before a certain amount of time.
\item In case reaching expiration time without the owner properly closing the channel,
  the user can close the channel receiving in return unexpended amount.
\end{itemize}

@@

Both components were heavily manually tested, during development
in order to assure the correct work-flows were correctly supported.

@@

To help developers and make technology more mature,
analysis tools\cite{ConsenSys19} were considered.

@@

For this project, Mythril from ConsenSys was used. Mythril is a
security analysis tool for EVM (Ethereum Virtual Machine) byte-code.
It detects security vulnerabilities in smart contracts built for Ethereum.

@@

Analysis was carried out over the Solidity-based smart contract called
StateChannel.sol.

@@

https://github.com/ethereum-internet-access/mythril
Firstly, we analyzed the local source code:

@@

Secondly, we analyzed on-chain smart contract.
Because the smart contract was deployed on Ropsten test-net we executed:

@@

To do so, simply install Docker and run the container over Raspbian/ARM.
It would deploy project's software. The script to execute is as follows:

@@

\begin{itemize}
\item Add firewall rules to prevent IP traffic between
  different end users.
\item Obviously safe the connectivity to the Raspberry
  using a wire connection and filtering any traffic on 22 (ssh)
  port coming from the wireless interface.
\item For this PoC, the captive portal IP address was used; because of that,
  TLS was not supported; accordingly the user had no guarantee of
  authenticity of the WiFi service provider.
\end{itemize}

@@

All smart contract-blockchain transactions are being processed
using Infura as an Ethereum proxy; main potential
security issue regarding
private key management arises when considering the ephemeral
private keypair generation on the front-end. Since this key
is not managed by MetaMask is not isolated from browser's
JavaScript engine being vulnerable to plain HTTP malicious
attacks by third parties and cross-site scripting issues.

@@

Apart from issues considered previously, general considerations
regarding safety of the system should take into account
the amount of value implied in the transactions;
being these micro-payments and not big amounts of value nor
big investments either, system can be considered relatively
safe in proportion to the assets (\textit{small Ethereum amounts})
transacted.

@@

Any organization that has customers on the premises,
from coffee shops to libraries, might want to offer free (or even
paid-for) Wi-Fi as an attraction. Collecting and storing personal
information brings responsibilities under European Union’s General
Data Protection Directive (GDPR). The GDPR\cite{gdpr16}, which took
effect on 25 May 2018, applies to any organization that handles EU
residents’ personal data. From 25 May on, however, organizations
failing to comply with the regulation face fines of up to 4 \% of
their global turnover. There are Certified EU General Data Protection
Regulation Foundation (GDPR) Training Courses.

@@

Another look at the major obligations imposed by this
regulation can be considered\cite{tata17}:

\vspace{0.35cm}
1. ART 5/1B and 1C - DATA MINIMIZATION:
Prohibited is the collection of more data than that required for the
adequate purposes for which it is being processed. For instance, if
you offer restaurant guests free WiFi access it is not necessary to
ask for their address data in this connection. However, if you are
interested in finding out where your guests are from - which can be a
completely legitimate interest for a restaurant owner to have - you
are indeed entitled to pose the question, but your guests must divulge
the information consciously and voluntarily and perhaps you should
only ask them for their postcode.

@@

\vspace{0.35cm}
5. ART 5/1F and 28 - INTEGRITY AND CONFIDENTIALITY: You must ensure
appropriate security of data and also ensure that it cannot be
tampered with. But how do you do that? What if hackers launch a
malicious attack? Don’t worry; even government websites are not immune
to this sort of thing and it would be hardly fair if the GDPR were to
impose more stringent requirements on you. However, you are required
to put appropriate technical or organizational measures in place that
provide adequate protection of your customers’ data. The GDPR does not
spell out in so many words what these required ‘measures’
are. However, you can be certain that you will be at least be expected
to have a valid SSL certificate for your data-collecting website and
to conclude instruction and confidentiality agreements with your
personnel who come into contact with the corresponding data.

@@

In addition, the Proposal for a new ePrivacy
Regulation was tabled in January 2017, when GDPR was not yet
applicable and the European Electronic Communications Code\cite{eecc18}
was still being negotiated. Now, GDPR has been implemented,
changing many of the approaches and behaviors that the ePrivacy
Regulation aims to tackle. Additionally, the new EECC already ensures
an extended definition of what an e-communications service is, thereby
achieving one of the purported goals of the ePrivacy Regulation, which
is to broaden the scope of application of confidentiality obligations
to a larger number of Internet-based players.

\vspace{0.35cm} Because of these issues, good advice for anyone
planning to set up a hot-spot would be to collect as little personal
data from users as possible, and to think of any additional collection
as a potential liability.

@@

In the case of 5G technologies for practical reasons
and the high density of antennas required,
different operators shall be able
to dynamically share the same resources (i.e. neutral host
infrastructure)\cite{neutral}. At the same time
in order to manage
service operation efficiently through its whole
life-cycle there's the need to automate as much as
possible all the work-flows; this requirements
perfectly fit smart contract capabilities
to model and operate over a defined set of rules
supporting payments, escrows, rewards or penalties
depending on different conditions.

\end{multicols}
\begin{center}
%\begin{figure}[h]
  %  \centering
  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/slices-y.eps}
%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
%\end{figure}
\\
Figure 6. Slicing model.
\\
\end{center}
\begin{multicols}{2}

As an example, in figure 6 it is depicted how three virtual operators over
different infrastructures shared or not in a slice-model of a neutral host
infrastructure provide their service;
a slice can be described as a dynamic multi-tenant virtualization
of some physical resources like access points,
physical switches, routers and data-centers including mobile edge computing
appliances; this is the case of currently ongoing projects like
5GCity\cite{5gcity} where a distributed cloud
and radio platform for municipalities and infrastructure owners acting
as 5G neutral hosts is being developed and tested. Platforms
targeting 5G business models such as 5GCity can benefit from adding
this smart contract business logic in order to bill end-users and automate
all the accounting layer.

@@

Although this is a promising approach; there's an issue with it regarding
regulation, in particular in the European Union, there's the requirement
that a telecommunication service provider must know their customers; this
is required for example to allow court requirements of users identity
in case there's some investigation of illegal activities. As a single payment
system, state channel smart contracts are pseudo-anonymous hence do not
support the \textit{Know Your Customer - KyC} required by regulation.

@@

To take into account the problem illustrated in previous paragraph
we can take as a basis a model of self-sovereign identity\cite{allen}
and in particular the one recently announced by the authorities
of Catalonia's government\cite{identicat} that will provide credentials
to citizens using DLT technologies as a Public Key Infrastructure; this
model could allow in some near future to give the citizen the option
of registering identity in such a way that her/his credentials can be linked
to some certified database that could be queried only by a judge
requirement and guarantee the users identities are recognized
%% and at the
%% same time hidden from service providers, discharging them from
%% any GDPR responsibilities apart from the intrinsic service provision
%% metadata.

@@

\bibitem{neutral} Enabling Neutral Host
Network Economics: CCS case study, GSM Association, August, 2018
\end{thebibliography}
\clearpage
